---
title: "Data Science Homework 1"
author: "Charlotte Fowler"
date: "9/12/2019"
output: github_document
---

```{r}
library(tidyverse)
```

# Problem 1

```{r}
#First, I the data frame as described in problem 1, with 4 variables.
data_frame_p1 = tibble(
  normal_sample = rnorm(8),
  normal_sample_positive = normal_sample > 0,
  vec_char = c("I", "need", "eight", "words", "to", "fill", "the", "vector"),
  vec_factor = factor(c("teal", "black", "teal", "white", "white", "teal", "black", "teal"))
)


#Now, to attempt taking the mean of each variable: 

#mean of normal sample
mean(pull(data_frame_p1, normal_sample))

#mean of logical vector
mean(pull(data_frame_p1, (normal_sample_positive)))

#mean of character vector
mean(pull(data_frame_p1, (vec_char)))

#mean of categorical vector
mean(pull(data_frame_p1, (vec_factor)))
```
I can take the mean of the sample values and of the logical vector as these  both contain have numerical values, however I cannot take the mean of the character nor categorical vector. 


```{r, results= 'hide', warning= FALSE}
#Applying 'as.numeric' function to logical, character, and categorical vectors
as.numeric(pull(data_frame_p1, (normal_sample_positive)))
as.numeric(pull(data_frame_p1, (vec_char)))
as.numeric(pull(data_frame_p1, (vec_factor)))
```
When applying the 'as.numeric' function to the logical, character, and categorical vectors, it is not possible with the character vector. This is becauase this vector contains abritrary elements, which R is not able to process as anything numerical. However, in the logical vector, R can interpret the 1 and 0 entries as plain numbers, and in the categorical vector R can assign numerical values 1, 2, and 3 to the three categories, alfabetically. This demonstrates how it's possible to take the mean of the logical vector above, as it has 1 and 0 values associated with True and False, respectively. Additionally, it shows how the character vector couldn't have the mean calculated above, as it is understood to be unrelated elements. With the categorical vector, if we had used the 'as.numeric' function above before taking the mean, we would have been able to, as R has now assigned a number to each category. 

```{r}
#logical vector to numeric and multiplied
as.numeric(pull(data_frame_p1, (normal_sample_positive))) * 
  pull(data_frame_p1, (normal_sample))
#logical vector to factor, then multipled
factor(pull(data_frame_p1, (normal_sample_positive))) * 
  pull(data_frame_p1, (normal_sample))
#logical vector to factor, then numeric, then multiplied
as.numeric(factor(pull(data_frame_p1, (normal_sample_positive)))) * 
  pull(data_frame_p1, (normal_sample))
```

